Cluster Dynamics with HETDEX - I: Simulated Performance, Mass Distribution and Limits

### 2016-01-21 ms.tex
added some text and tweaked
should now be working pretty well.

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
% \usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
% \usepackage{mathptmx}
% \usepackage{txfonts}

@@

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
\input{commands.tex}

% Only include extra packages if you really need them. Common packages are:
%\usepackage{epsf}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{natbib}
%\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}
\usepackage{etoolbox}

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\newcommand{\editorial}[1]{\textcolor{red}{#1}}
\DeclareRobustCommand{\ion}[2]{%
\relax\ifmmode
\ifx\testbx\f@series
{\mathbf{#1\,\mathsc{#2}}}\else
{\mathrm{#1\,\mathsc{#2}}}\fi
\else\textup{#1\,{\mdseries\textsc{#2}}}%
\fi}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. Boada et al.]{
Steven Boada,$^{1}$\thanks{E-mail: boada@physics.tamu.edu (SB)}
C.~Papovich,$^{1}$
R.~Wechsler$^{2,3}$
\\
% List of institutions
$^{1}$George P.\ and Cynthia Woods Mitchell Institute for
Fundamental Physics and Astronomy, Texas A\&M University, College Station, TX, 77843-4242\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}

@@

\begin{abstract}
\noindent
ABSTRACT GOES HERE!!
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain, as best possible, $P(X|M,z)$ the probability ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using observable parameter ($X$). The observable parameter most commonly observed X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014, Mantz2015}, cosmic microwave background observations \citeeg{Vanderlinde2010, Sehgal2011} using the Sunyaev--Zel’dovich effect (SZE; \citealt{Sunyaev1972}) or optical studies \citeeg{Rozo2010, Rozo2015} of richness \citeeg{Abell1958, Rykoff2012} or galaxy velocity dispersions \citeeg{Ruel2014, Sifon2015}.

@@

\begin{figure} 
	\includegraphics[width=0.49\textwidth]{hmf.pdf} 
	\caption{The cumulative MF of halos above $M_{200c}$ at $z=0.1$. The predicted MF is from tinker2008.} 
	\label{fig: hmf} 
\end{figure}

@@

\subsection{ {\rm[\ion{O}{ii}]} Luminosity}\label{sec: oii luminosity}
\begin{figure*} 
	\includegraphics[width=\textwidth]{oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. New [\ion{O}{ii}] luminosity (and subsequently fluxes) are assigned to catalog galaxies from slice sampling the probability histogram.} \label{fig: oii sdss} 
\end{figure*}

@@

\begin{figure} 
	\includegraphics[width=0.49\textwidth]{f01.pdf} 
	\caption{Representative observation tiling scheme for the HETDEX $16' \times 16'$ pointings. Each colored square is a single VIRUS IFU and the dashed octagons approximate the size of a single observation. See the text for more details.} \label{fig: ifu layout} 
\end{figure}

@@

\subsubsection{Probability Based}
We begin with the training sample. After selecting the desired features $\vec{x} = \{\sigma, z, ...\}$ we make the joint probability between the true cluster mass (M) and $\vec{x}$. Because $\vec{x}$ can be multidimensional, we rely on the corner plot to visualize the relationship between all of the training features. Figure~\ref{fig: probability corner} shows all of the one (marginalized probability) and two (joint probability) dimensional projections of the posterior probability distributions of the features of the training data.

\begin{figure} 
	\includegraphics[width=0.5\textwidth]{cornertest.pdf} 
	\caption{Corner plot of the \emph{training} data with features $\sigma$ and $z$. The corner plots shows all of the one and two dimensional posterior probability distributions used to determine the correct cluster mass. The colored rectangles show the slices needed to create a conditional probability distribution of the mass, $P(M|\vec{x})$. See text for a complete description. } \label{fig: probability corner} 
\end{figure}

@@

\subsubsection{Machine Learning Based}
The estimation in this section relies on a ML technique known as an ensemble method, where many estimators are created by a single learning method with the goal of improved generalization and robustness compared to a single estimation. Ensemble methods come in two general flavors. Averaging methods average (hence the name) the estimators to produce a single prediction. Boosting estimators build estimates sequentially by attempting to address poor performing estimators in each previous step, hence ``boosting'' the predictive power.

Here we use an averaging ensemble learning method known as a forest of randomized decision trees often shorten to just random forest (RF). Decision trees can be visualized a flow chart where forks are the branches of the tree. The path along the tree is decided by the values of the feature at each branch. RF estimators use a random subset of the training set at each fork to decide which path should be followed. The final prediction is then the average of all the trees. We use RF regression methods as implemented in {\sc Scikit-Learn} \citep{Pedregosa2012}.

@@

\begin{figure*} 
	\includegraphics[width=\textwidth]{recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and mass for the targeted (upper row) and survey (lower row) observing strategies. The significant decline in galaxies observed with the survey strategy is due to gaps in the VIRUS IFU.} \label{fig: recovery} 
\end{figure*}

@@

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{MLcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq: mass scaling}) and the ML technique with different input features. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) as a function of mass. The solid black line shows the 1:1 relation. The colored lines and shaded region show the median value and 68\% bounds respectively.} \label{fig: ML comparison} 
\end{figure*}

\begin{equation}\label{eq: fractional error}
	\epsilon = (M_{pred} - M)/M
\end{equation}

\editorial{It doesn't look like you can use the targeted training sample to train the survey data. If you do, it either does worse or about the same if you didn't use any training at all. After looking at it, you do have to use the survey training data to get good results about 0.13 dex MAE
}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

@@

\section*{Acknowledgements}
The authors also wish to thank the anonymous referee whose comments and suggestions significantly improved both the quality and clarity of this work. We also thank Steven W. Crawford for many helpful discussions. This research made use of This research made use of \textsc{APLpy}, an open-source plotting package for Python hosted at http://aplpy.github.com; the \textsc{IPython} package \citep{Perez2007}; \textsc{matplotlib}, a Python library for publication quality graphics \citep{Hunter2007}. Funding for the SDSS and SDSS-II has been provided by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S. Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England. The SDSS Web Site is http://www.sdss.org/. The SDSS is managed by the Astrophysical Research Consortium for the Participating Institutions.

>>>>

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
% \usepackage{newtxtext,newtxmath}
\usepackage{times}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
% \usepackage{mathptmx}
% \usepackage{txfonts}

@@

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
%\usepackage{epsf}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{natbib}
%\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}
\usepackage{etoolbox}

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\input{commands.tex}

\newcommand{\editorial}[1]{\textcolor{red}{#1}}
\DeclareRobustCommand{\ion}[2]{%
\relax\ifmmode
\ifx\testbx\f@series
{\mathbf{#1\,\mathsc{#2}}}\else
{\mathrm{#1\,\mathsc{#2}}}\fi
\else\textup{#1\,{\mdseries\textsc{#2}}}%
\fi}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. Boada et al.]{
Steven Boada,$^{1}$\thanks{E-mail: boada@physics.tamu.edu}
C.~Papovich,$^{1}$
R.~Wechsler,$^{2,3}$
T. S.~Li,$^{1}$\newauthor\phantom{x}
K.~Gebhardt,$^{4}$
E.~Rozo,$^{2}$
E. S.~Rykoff,$^{2}$
\\
% List of institutions
$^{1}$George P.\ and Cynthia Woods Mitchell Institute for
Fundamental Physics and Astronomy, Texas A\&M University, College Station, TX, 77843-4242\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA\\
$^{4}$The University of Texas}

@@

\begin{abstract}
\noindent
The study of clusters of galaxies has been argued to be a very effective way to measure cosmological parameters, including measuring dark energy and testing models of gravity. The Hobby Eberly Telescope Dark Energy Experiment (HETDEX) will observe many hundreds of square degrees, covering a large sample of galaxy clusters out to $z = 0.5$ based on their optical spectra ($3500-5500\AAA$). The spectra will provide important measures of the clusters dynamics and may enable constraints on cosmological parameters, but only if the measurements provide accurate estimates of the total cluster masses. We have carried out a study to investigate the ability of HETDEX to recover accurate galaxy cluster masses over a wide range of masses and redshifts. We used a detailed mock galaxy catalog and present mock observations of two different scenarios: (1) We targeted individual galaxy clusters to investigate the recovery of parameters with such observations. (2) We created and evaluated a HETDEX-like selection "function'' of galaxies over a similarly sized portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and mass. Using both observing strategies, we produce cluster mass probability density functions $P(X|M,z)$, which can be used to determine the probability that a galaxy cluster of given mass (M), located at redshift ($z$) determined using observable parameter (X). 
% We then applied these probability functions to ten galaxy clusters selected from the Sloan Digital Sky Survey DR8 and the Chandra-XMM X-ray Cluster Survey at $z=0.2-0.3$, and observed by the HETDEX spectrograph prototype instrument (VIRUS-p). We measured spectroscopic redshifts and line-of-sight velocities of the galaxies in and around each cluster, derived a line-of-sight velocity dispersion, and inferred a dynamical mass for each cluster which ranges from $(0.4-24) \times 10^{14} \Msol (M_{200c})$. Using the mass probability density functions described above, we updated these masses and compared them to existing literature estimates, or to the masses estimated from other observables such as X-ray temperature or richness.
\end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain, as best possible, $P(X|M,z)$ the probability ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using observable parameter ($X$). The observable parameter most commonly observed X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014, Mantz2015}, cosmic microwave background observations \citeeg{Vanderlinde2010, Sehgal2011} using the Sunyaev-Zel'dovich effect (SZE; \citealt{Sunyaev1972}) or optical studies \citeeg{Rozo2010, Rozo2015} of richness \citeeg{Abell1958, Rykoff2012} or galaxy velocity dispersions \citeeg{Ruel2014, Sifon2015}.

@@

\begin{figure} 
	\includegraphics[width=\columnwidth]{figures/hmf.pdf} 
	\caption{The cumulative MF (from \citealt{Tinker2008}) of halos above $M_{200c}$ at $z=0.1$} 
	\label{fig: hmf} 
\end{figure}

@@

\subsection{ {\rm[\ion{O}{ii}]} Luminosity}\label{sec: oii luminosity}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. New [\ion{O}{ii}] luminosity (and subsequently fluxes) are assigned to catalog galaxies from slice sampling the probability histogram.} \label{fig: oii sdss} 
\end{figure*}

@@

\begin{figure} 
	\includegraphics[width=\columnwidth]{figures/f01.pdf} 
	\caption{Representative observation tiling scheme for the HETDEX $16' \times 16'$ pointings. Each colored square is a single VIRUS IFU and the dashed octagons approximate the size of a single observation. See the text for more details.} \label{fig: ifu layout} 
\end{figure}

@@

\subsubsection{Probability Based}\label{sec:probability method}
We begin with the training sample. After selecting the desired features $\vec{x} = \{\sigma, z, ...\}$ we make the joint probability between the true cluster mass (M) and $\vec{x}$. Because $\vec{x}$ can be multidimensional, we rely on the corner plot to visualize the relationship between all of the training features. Figure~\ref{fig: probability corner} shows all of the one (marginalized probability) and two (joint probability) dimensional projections of the posterior probability distributions of the features of the training data.

\begin{figure} 
	\includegraphics[width=\columnwidth]{figures/cornertest.pdf} 
	\caption{Corner plot of the \emph{training} data with features $\sigma$ and $z$. The corner plots shows all of the one and two dimensional posterior probability distributions used to determine the correct cluster mass. The colored rectangles show the slices needed to create a conditional probability distribution of the mass, $P(M|\vec{x})$. See text for a complete description. } \label{fig: probability corner} 
\end{figure}

@@

\subsubsection{Machine Learning Based}\label{sec:machine learning method}
The estimation in this section relies on a ML technique known as an ensemble method, where many estimators are created by a single learning method with the goal of improved generalization and robustness compared to a single estimation. Ensemble methods come in two general flavors. Averaging methods average (hence the name) the estimators to produce a single prediction. Boosting estimators build estimates sequentially by attempting to address poor performing estimators in each previous step, hence ``boosting'' the predictive power.

Here we use an averaging ensemble learning method known as a forest of randomized decision trees often shorten to just random forest (RF). Decision trees can be visualized a flow chart where forks are the branches of the tree. The path along the tree is decided by the values of the feature at each branch. RF estimators use a random subset of the training set at each fork to decide which path should be followed. The final prediction is then the average of all the trees. We use RF regression methods as implemented in {\tt Scikit-Learn} \citep{Pedregosa2012}.

@@

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and mass for the targeted (upper row) and survey (lower row) observing strategies. The significant decline in galaxies observed with the survey strategy is due to gaps in the VIRUS IFU.} \label{fig: recovery} 
\end{figure*}

@@

\subsection{Mass estimates}
In this section we discuss the how accurately we are able to reproduce the true cluster mass from a set of observations. We report on two methods the probability based approach (Section \ref{sec:probability method}) and the machine learning based method (Section \ref{sec:machine learning method}). For each method we consider both targeted and HETDEX-like observing strategies.

All of the figures presented here show the predicted versus true cluster masses for one of the two observing strategies. In each panel the solid black line shows the 1:1 relationship, the colored, solid line is the median recovered mass, and the shaded region is the 68\% scatter of the cluster masses. The lower panels show the fractional mass error: 
\begin{equation}\label{eq: fractional error}
	\epsilon = (M_{pred} - M)/M
\end{equation}
where $M_{pred}$ is the predicted cluster mass and $M$ is the true cluster mass.

\subsubsection{Probability Based}
Figures \ref{fig:Probability comparison_survey} and \ref{fig:Probability comparison} show the power law predictions compared to the predictions based on probability propagation.

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison_survey.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq: mass scaling}) and the ML technique with different input features. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) as a function of mass. The solid black line shows the 1:1 relation. The colored lines and shaded region show the median value and 68\% bounds respectively.} \label{fig:Probability comparison_survey} 
\end{figure*}

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq: mass scaling}) and the ML technique with different input features. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) as a function of mass. The solid black line shows the 1:1 relation. The colored lines and shaded region show the median value and 68\% bounds respectively.} \label{fig:Probability comparison} 
\end{figure*}

\subsubsection{Machine Learning Based}
Figures \ref{fig: ML comparison_survey} and \ref{fig: ML comparison} show the power law predictions compared to the RF predictions using LOSVD ($\sigma$) combined with both $z$ and the number of galaxies observed, $N_{gal}$.

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison_survey.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq: mass scaling}) and the ML technique with different input features. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) as a function of mass. The solid black line shows the 1:1 relation. The colored lines and shaded region show the median value and 68\% bounds respectively.} \label{fig: ML comparison_survey} 
\end{figure*}

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption{Mass predictions for the power law scaling relation (Equation~\ref{eq: mass scaling}) and the ML technique with different input features. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) as a function of mass. The solid black line shows the 1:1 relation. The colored lines and shaded region show the median value and 68\% bounds respectively.} \label{fig: ML comparison} 
\end{figure*}

% \editorial{It doesn't look like you can use the targeted training sample to train the survey data. If you do, it either does worse or about the same if you didn't use any training at all. After looking at it, you do have to use the survey training data to get good results about 0.13 dex MAE
% }

@@

\section*{Acknowledgements}
The authors also wish to thank the anonymous referee whose comments and suggestions significantly improved both the quality and clarity of this work. We also thank Steven W. Crawford for many helpful discussions. This research made use of This research made use of  the {\tt IPython} package \citep{Perez2007} and {\tt matplotlib}, a Python library for publication quality graphics \citep{Hunter2007}. Funding for the SDSS and SDSS-II has been provided by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S. Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England. The SDSS Web Site is http://www.sdss.org/. The SDSS is managed by the Astrophysical Research Consortium for the Participating Institutions.
